---
title: "Inference Exercise"
author: "Keith Hultman"
date: "October 10, 2016"
output: html_document
---

# Using dplyr and creating histograms of data. 


## Analysis of heights

Let's recreate the histogram on heights from last week's lecture. Import the dataset and examine it. 
```{r definingHeights, message=FALSE}
data(father.son,package="UsingR")
head(father.son)

```

Select the father heights column (fheights) and place into the variable 'father'. We will be using the dplyr package for data manipulation. This package uses 'pipes' (%>%) to send the data through several functions. First we have our variable and assignment operator (<-) followed by the data frame. Then we have a series of functions connected by pipes. In this case we just select one column called fheight, then we have to use the unlist function to create a vector from the data table. 

And then take the mean of the father's heights.


```{r}
library(dplyr)
father <- father.son %>% select(fheight) %>% unlist
mean(father)
```

If we wanted to filter the data we could include a filter function. The following returns the father heights when the son's height is taller than the father average.

```{r}
filter_example <- father.son %>%  
  filter(sheight > mean(father)) %>% 
  select(fheight) %>% 
  unlist

mean(filter_example)
```

## Histogram
Create a histogram of the heights of all fathers in the original data

```{r hist, eval=FALSE}
hist(father)
```


# Analysis of life expectancy

Do the same analysis for life expectancy data. First we will extract data from the gapminder dataset, which includes data on life expectancy and gdp for several countries over many years. 

```{r}
library(gapminder)
data(gapminder)
head(gapminder)
```

Create a vector 'y1952' of the life expectancies of each country for the year 1952, using the filter(year == 1952) function.

```{r}

```


Plot a histogram of these life expectancies to see the spread of the different countries.

```{r}

```

Make a histogram of life expectancies for the most recent year in the dataset, 2007.

```{r}

```

